Global adjoint tomography -- Model GLAD-M25\footnote{This paper is dedicated to Dimitri Komatitsch, computational geophysicist extraordinaire, who passed away unexpectedly shortly before its submission.}

### 2019-02-15 main.tex
Updates from Overleaf

In addition to the 253 earthquakes used for the construction of GLAD-M15, we initially carefully selected 784 additional earthquakes
from the global Centroid-Moment Tensor (CMT) catalog~\citep[e.g.,][]{ekstrom2012global},
resulting in a total number of 1,040 earthquakes.
To ensure a good signal-to-noise ratio on a global scale,
the smallest moment magnitude in the database is set to~5.5,
thereby capturing more deep events,
and to avoid complications associated with source dimension and directivity,
the largest moment magnitude is set to~7.2.

@@

The seismographic stations used in this study were carefully selected to ensure global coverage
and high data quality (Fig.~\ref{fig:stations}).
In addition to the seismographic  stations used for the source inversions,
we included all available data for our 1,480 event earthquake database from many data centers,
including IRIS, ORFEUS, INGV, IPGP, ETH, and GEONET.
Regional and temporary networks,
such as US Array~(TA),
Africa Array~(AF), the Canadian National Seismograph Network~(CN), Geoscience Australia~(AU),
the Antarctic Seismographic Argentinean Italian Network~(AI),
and the New Zealand National Seismograph Network~(NZ),
constitute a significant part
of our database and greatly improve coverage in certain regions. \red{(Looks like there are also OBSs to be mentioned, mainly for my curiosity, it would be good to know their contribution to the overall inversions.)}

@@

Conventional seismic data formats, such as SAC, involve one file per time series
plus related files with instrument response and event information.
Since every earthquake in the database is typically recorded by thousands of
instruments, I/O during the preprocessing stage of the adjoint tomography workflow
quickly cripples the file system. The Adaptable Seismic Data Format
(ASDF)~\citep[][ \texttt{https://seismic-data.org/ }]{krischer2016adaptable}
was developed with complete reproducibility and fast parallel processing in mind.

@@

To illustrate which seismic phases end up being picked,
Figs.~\ref{fig:window_density_Z}, \ref{fig:window_density_R}, and \ref{fig:window_density_T} show time-distance plots
for vertical, radial, and transverse component 17--40~s body waves,
respectively,
in which the Pyflex window count is used to identify parts of seismograms being assimilated.
The figures nicely highlight all the well known body traveltime branches.
Collectively, the numerous compressional wave arrivals identified in Figs.~\ref{fig:window_density_Z} and~\ref{fig:window_density_R} are of great importance, because they help constrain the compressional wavespeed structure in the model.
The Pyflex window selection algorithm is currently configured not to pick body waves after the arrival of the surface waves, which is why the PcP and ScS branches are missing at shorter epicentral distances \red{(Again just for my curiosity, why PcP and ScS phases together with any phase after surface waves were excluded? Any special reason or just a decision?)}.
This is something we will reconsider in future applications, perhaps by introducing aspects of machine learning.

@@

\begin{figure}
  \centering
  \includegraphics[width=0.9\textwidth]{figures/depth_slice/globe_vs.pdf}
  \caption{\small{Map views of global shear wavespeed perturbations at various depths for  model
  GLAD-M25~(left column), S362ANI$+$M~\citep[middle column;][]{moulik2014anisotropic},
  and S40RTS~\citep[right column;][]{ritsema2011s40rts}.
  At a given radius,
  perturbations are calculated relative to each model's average.
  The range of perturbations in shear wavespeed varies from map to map, as indicated in the top left of each panel.
  The green circles denote locations of various
  hotspots~\citep{montelli2006catalogue}.
  The range of the colorbar is the same for each row,
  and its maximum value is indicated in parentheses on the left, after
  the depth. \red{(Probably the numbers on top-left of each map are min/max perturbations of the associated sections, but the colorbar is fixed to 2 to 7$\%$)}}}
  \label{fig:global-vs}
\end{figure}

@@

\begin{figure}[ht!] 
  \includegraphics[width=0.9\textwidth]{figures/depth_slice/south_america_vp.pdf}
  \caption{\small{Map views of compressional wavespeed variations beneath South America for GLAD-M25~(first column) and global models S362ANI$+$M~\citep{moulik2014anisotropic} (second column), LLNL-G3Dv3~\citep{simmons2012llnl} (third column), and GAP-P4~\citep{fukao2013subducted} (fourth column).\red{(LLNL model looks pretty interesting with the craton signature down to 300 km. Slab and craton anomalies together at 150 km look interesting too, as if they are two different models. In addition I am still wondering why are we the only one seeing Scotia Arc? Or why the others do not see it? It looks to me that it is pretty easy to add a slab signature to Scotia because of USArray. If there are earthquakes in their database at Scotia it should still easy to put it there even without USArray. Or perhaps it is indeed the bias of USArray coming from GLAD-M15?.)}}}
  \label{fig:southamerica-vp}
  \centering
\end{figure}

@@

\begin{figure}[ht!]
    \centering
    \sidesubfloat[]{\includegraphics[width=0.98\textwidth]{figures/plumes/Afar.png}}\\[-1pt]
    \sidesubfloat[]{\includegraphics[width=0.98\textwidth]{figures/plumes/Bermuda_Canary.png}}\\[-1pt]
    \sidesubfloat[]{\includegraphics[width=0.98\textwidth]{figures/plumes/CapeVerde_Hoggar.png}}\\[-1pt]
    \sidesubfloat[]{\includegraphics[width=0.98\textwidth]{figures/plumes/Iceland_Eifel.png}}\\[-1pt]
    \sidesubfloat[]{\includegraphics[width=0.98\textwidth]{figures/plumes/Easter_Galapagos.png}}\\[-1pt]
    \sidesubfloat[]{\includegraphics[width=0.98\textwidth]{figures/plumes/Marion3_Kerguelen.png}}\\
    \caption{\small{Vertical cross sections for various plumes for models GLAD-M25 ($V_\textrm{S}$; left column), TX2015~\citep[$V_\textrm{S}$;][]{TX2015} (center column), and SEMUCB-WM1~\citep[$V_\textrm{S}$;][]{french2015broad} (right column).
    The map in the top left corner of each row shows the cross section with color-coded red, green, and white dots for geographical reference;
    hotspots are denoted by red triangles.
    The dashed black semicircles in the cross sections denote depths of 410~km, 660~km, and 1000~km.
    (a) Afar; (b) Bermuda (left) and Canary (middle); (c) Cape Verde (middle) and Hoggar (right); (d) Iceland (middle) and Eifel (right); (e) Easter (left) and Galapagos (right); (f) Marion (middle), and Kerguelen (right). \red{(GLAD-M25 is still Vp)}}}
    \label{fig:plumes}
\end{figure}

In this section we highlight various plume systems on Earth. Fig.~\ref{fig:plumes}
shows cross sections of several plumes, comparing models GLAD-M25 ($V_\textrm{P}$),
TX2015~\citep[$V_\textrm{S}$;][]{TX2015}, and SEMUCB-WM1~\citep[$V_\textrm{S}$;][]{french2015broad}.
All three models have common plume features where GLAD-M25 and TX2015 seem to be more in good agreement.

@@

\begin{figure}
    \centering
    \sidesubfloat[]{\includegraphics[width=0.88\textwidth]{figures/subductions/Aeg.png}\label{fig:subd_a}}\\[-8pt]
    \sidesubfloat[]{\includegraphics[width=0.88\textwidth]{figures/subductions/Br.png}\label{fig:subd_b}}\\[-8pt]
    \sidesubfloat[]{\includegraphics[width=0.88\textwidth]{figures/subductions/Ha.png}\label{fig:subd_c}}\\[-8pt]
    \sidesubfloat[]{\includegraphics[width=0.88\textwidth]{figures/subductions/Wc.png}\label{fig:subd_e}}\\[-8pt]
    \sidesubfloat[]{\includegraphics[width=0.88\textwidth]{figures/subductions/Ne.png}\label{fig:subd_d}}\\[-8pt]
    \sidesubfloat[]{\includegraphics[width=0.88\textwidth]{figures/subductions/Su.png}\label{fig:subd_f}}\\[-8pt]
    \caption{\small{Vertical cross sections of compressional wavespeed perturbations in various subduction zones
    for models GLAD-M25 ($V_\textrm{P}$; left column), GAP-P4~\citep{fukao2013subducted} (middle column) and UU-P07~\citep{van2018atlas} (right column).
    The map in the top left corner of each row shows the cross section with color-coded red, green, and white dots for geographical reference.
    The dashed black semicircles in the cross sections denote depths of 410~km, 660~km, and 1000~km.
    (a) Aegean; (b) South America; (c) Hatteras; (d) Wichita. (e) Nepal; (f) Sunda.}}
    \label{fig:subd}
\end{figure}

@@

Seismology is a science driven by data.
There was a time when seismologists would meticulously inspect single seismograms to glean valuable information from them \red{(Perhaps it is now the case for planetary seismology? )}.
Today, we have easy access to data from thousands of earthquakes recorded by tens of thousands of seismographic stations, and individual inspection of all of these records has become impossible \red{(probably it depends on the seismologist. I remember Jeroen Ritsema telling that it is not a big deal to go through the entire database in a week or so with some not-so-sophisticated plotting techniques. Those interested in data still look at, I think..)}.
Still, it would be a shame not to utilize as much as possible of the information in this massive seismographic database.

>>>>

In addition to the 253 earthquakes used for the construction of GLAD-M15, we initially carefully selected 784 additional earthquakes
from the global Centroid-Moment Tensor (CMT) catalog~\citep[e.g.,][]{ekstrom2012global},
resulting in a total number of 1,040 earthquakes.
To ensure a good signal-to-noise ratio on a global scale,
the smallest moment magnitude in the database is set to~5.5,
thereby capturing more deep events,
and, to avoid complications associated with source dimension and directivity,
the largest moment magnitude is set to~7.2.

@@

The seismographic stations used in this study were carefully selected to ensure global coverage
and high data quality (Fig.~\ref{fig:stations}).
In addition to the seismographic  stations used for the source inversions,
we included all available data for our 1,480 event earthquake database from many data centers,
including IRIS, ORFEUS, INGV, IPGP, ETH, and GEONET.
Regional and temporary networks,
such as US Array~(TA),
Africa Array~(AF), the Canadian National Seismograph Network~(CN), Geoscience Australia~(AU),
the Antarctic Seismographic Argentinean Italian Network~(AI),
and the New Zealand National Seismograph Network~(NZ),
constitute a significant part
of our database and greatly improve coverage in certain regions.

@@

Conventional seismic data formats, such as SAC, involve one file per time series
plus related files with instrument response and event information.
Since every earthquake in the database is typically recorded by thousands of
instruments, I/O during the preprocessing stage of the adjoint tomography workflow
quickly cripples the file system. The Adaptable Seismic Data Format
(ASDF)~\citep[][ \texttt{https://seismic-data.org/}]{krischer2016adaptable}
was developed with complete reproducibility and fast parallel processing in mind.

@@

To illustrate which seismic phases end up being picked,
Figs.~\ref{fig:window_density_Z}, \ref{fig:window_density_R}, and \ref{fig:window_density_T} show time-distance plots
for vertical, radial, and transverse component 17--40~s body waves,
respectively,
in which the Pyflex window count is used to identify parts of seismograms being assimilated.
The figures nicely highlight all the well known body traveltime branches.
Collectively, the numerous compressional wave arrivals identified in Figs.~\ref{fig:window_density_Z} and~\ref{fig:window_density_R} are of great importance, because they help constrain the compressional wavespeed structure in the model.
The Pyflex window selection algorithm is currently configured not to pick body waves after the arrival of the surface waves, which is why the PcP and ScS branches are missing at shorter epicentral distances.
This is something we will reconsider in future applications, perhaps by introducing aspects of machine learning.

@@

\begin{figure}
  \centering
  \includegraphics[width=0.9\textwidth]{figures/depth_slice/globe_vs.pdf}
  \caption{\small{Map views of global shear wavespeed perturbations at various depths for  model
  GLAD-M25~(left column), S362ANI$+$M~\citep[middle column;][]{moulik2014anisotropic},
  and S40RTS~\citep[right column;][]{ritsema2011s40rts}.
  At a given radius,
  perturbations are calculated relative to each model's average.
  The range of perturbations in shear wavespeed varies from map to map, as indicated in the top left of each panel.
  The green circles denote locations of various
  hotspots~\citep{montelli2006catalogue}.
  The range of the colorbar is the same for each row,
  and its maximum value is indicated in parentheses on the left, after
  the depth.}}
  \label{fig:global-vs}
\end{figure}

@@

\begin{figure}[ht!] 
  \includegraphics[width=0.9\textwidth]{figures/depth_slice/south_america_vp.pdf}
  \caption{\small{Map views of compressional wavespeed variations beneath South America for GLAD-M25~(first column) and global models S362ANI$+$M~\citep{moulik2014anisotropic} (second column), LLNL-G3Dv3~\citep{simmons2012llnl} (third column), and GAP-P4~\citep{fukao2013subducted} (fourth column).}}
  \label{fig:southamerica-vp}
  \centering
\end{figure}

@@

\begin{figure}[ht!]
    \centering
    \sidesubfloat[]{\includegraphics[width=0.98\textwidth]{figures/plumes/Afar.png}}\\[-1pt]
    \sidesubfloat[]{\includegraphics[width=0.98\textwidth]{figures/plumes/Bermuda_Canary.png}}\\[-1pt]
    \sidesubfloat[]{\includegraphics[width=0.98\textwidth]{figures/plumes/CapeVerde_Hoggar.png}}\\[-1pt]
    \sidesubfloat[]{\includegraphics[width=0.98\textwidth]{figures/plumes/Iceland_Eifel.png}}\\[-1pt]
    \sidesubfloat[]{\includegraphics[width=0.98\textwidth]{figures/plumes/Easter_Galapagos.png}}\\[-1pt]
    \sidesubfloat[]{\includegraphics[width=0.98\textwidth]{figures/plumes/Marion3_Kerguelen.png}}\\
    \caption{\small{Vertical cross sections for various plumes for models GLAD-M25 ($V_\textrm{S}$; left column), TX2015~\citep[$V_\textrm{S}$;][]{TX2015} (center column), and SEMUCB-WM1~\citep[$V_\textrm{S}$;][]{french2015broad} (right column).
    The map in the top left corner of each row shows the cross section with color-coded red, green, and white dots for geographical reference;
    hotspots are denoted by red triangles.
    The dashed black semicircles in the cross sections denote depths of 410~km, 660~km, and 1000~km.
    (a) Afar; (b) Bermuda (left) and Canary (middle); (c) Cape Verde (middle) and Hoggar (right); (d) Iceland (middle) and Eifel (right); (e) Easter (left) and Galapagos (right); (f) Marion (middle), and Kerguelen (right). }}
    \label{fig:plumes}
\end{figure}

In this section we highlight various plume systems on Earth. Fig.~\ref{fig:plumes}
shows cross sections of several plumes, comparing models GLAD-M25 ($V_\textrm{P}$),
TX2015~\citep[$V_\textrm{S}$;][]{TX2015}, and SEMUCB-WM1~\citep[$V_\textrm{S}$;][]{french2015broad}.
All three models have common plume features,
and GLAD-M25 and TX2015 are in overall good agreement.

@@

\begin{figure}
    \centering
    \sidesubfloat[]{\includegraphics[width=0.88\textwidth]{figures/subductions/Aeg.png}\label{fig:subd_a}}\\[-8pt]
    \sidesubfloat[]{\includegraphics[width=0.88\textwidth]{figures/subductions/Br.png}\label{fig:subd_b}}\\[-8pt]
    \sidesubfloat[]{\includegraphics[width=0.88\textwidth]{figures/subductions/Ha.png}\label{fig:subd_c}}\\[-8pt]
    \sidesubfloat[]{\includegraphics[width=0.88\textwidth]{figures/subductions/Wc.png}\label{fig:subd_e}}\\[-8pt]
    \sidesubfloat[]{\includegraphics[width=0.88\textwidth]{figures/subductions/Ne.png}\label{fig:subd_d}}\\[-8pt]
    \sidesubfloat[]{\includegraphics[width=0.88\textwidth]{figures/subductions/Su.png}\label{fig:subd_f}}\\[-8pt]
    \caption{\small{Vertical cross sections of compressional wavespeed perturbations in various subduction zones
    for models GLAD-M25 ($V_\textrm{P}$; left column), GAP-P4~\citep{fukao2013subducted} (middle column) and UU-P07~\citep{van2018atlas} (right column).
    The map in the top left corner of each row shows the cross section with color-coded red, green, and white dots for geographical reference.
    The dashed black semicircles in the cross sections denote depths of 410~km, 660~km, and 1000~km.
    (a) Aegean; (b) South America; (c) Hatteras; (d) Wichita. (e) Nepal; (f) Sunda. }}
    \label{fig:subd}
\end{figure}

@@

Seismology is a science driven by data.
There was a time when global seismologists would meticulously inspect single seismograms to glean valuable information from them.
Today, we have easy access to data from thousands of earthquakes recorded by tens of thousands of seismographic stations, and individual inspection of all of these records has become impossible.
Still, it would be a shame not to utilize as much as possible of the information in this massive seismographic database.
